### 2020-08-28 Untitled.Rmd
methods almost fixed

---
title: "test"
author: "Dan"
date: "8/27/2020"
output:
  bookdown::pdf_document2: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown

Table.\@ref(tab:param-tab)

```{r param-tab}

cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "K", `Default Prior` = "$logn(10\\times{max(catch)},4)$") %>%
  rbind(c("Growth rate", "r", "$logn(r_{fishlife}, \\sigma_{r,fishlife})$")) %>%
  rbind(c("Shape parameter", "m", "Drawn from @horson2012")) %>%
  rbind(c("Catchability", "q", "$logn(1e^{-3}, 0.3)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,2)$")) %>%
  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.05,0.5)$")) %>%
  rbind(c("Tech Creep", "$creep$", "$logn(0.025,0.2)$")) %>% 
  rbind(c("Initial State", "$B0$", "$logn(0.025,0.2)$"))

knitr::kable(
  cd,
  "latex",
  align = "l",
  escape = F,
  row.names = F,
  caption = "Potential parameters included in sraplus, abbreviations, and default priors"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```