### 2021-06-26 documents/ovando-etal-assessing-global-fisheries.Rmd
merging final version back to main

---
title: Improving Estimates of the State of Global Fisheries <br> Depends on Better Data
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: fish-and-fisheries.csl
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
params:
  results_name: ["v1.0"]
  min_years_catch: [25]
always_allow_html: true
linkcolor: blue
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 600,
                      cache = FALSE, fig.width = 6, fig.asp = .75, dev = "png")
library(tidyverse)
library(viridis)
library(sf)
library(hrbrthemes)
library(extrafont)
library(scales)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(here)
library(sraplus)
extrafont::loadfonts()

functions <- list.files(here::here("functions"))

purrr::walk(functions, ~ source(here::here("functions", .x)))

results_name <- params$results_name

results_path <- here("results",results_name)

min_years_catch <- params$min_years_catch

prepare_sofia_data(lookup_fmi_names = FALSE)

ram_comp_data <- ram_data %>%
  mutate(has_things = !(is.na(catch) | catch == 0)) %>%
  filter(has_things) %>%
  group_by(stockid) %>%
  mutate(delta_year = as.integer(year - lag(year))) %>%
  mutate(delta_year = case_when(year == min(year) ~ as.integer(1),
                                TRUE ~ delta_year)) %>%
  mutate(missing_gaps = any(delta_year > 1)) %>%
  filter(missing_gaps == FALSE) %>%
  group_by(stockid) %>%
  mutate(n = length(catch)) %>%
  filter(n >= min_years_catch) %>%
  ungroup()

ram_comp_data <- ram_comp_data %>%
  select(
    year,
    stockid,
    country,
    primary_FAOarea,
    scientificname,
    commonname,
    year,
    catch,
    b_v_bmsy,
    u_v_umsy,
    total_biomass
  ) %>%
  rename(scientific_name = scientificname) %>% 
  mutate(fao_area_code = as.numeric(primary_FAOarea)) %>% 
  left_join(fao_species, by = "scientific_name")

assess_ram_fits <- read_rds(file = file.path(results_path,"assess_ram_fits.rds"))

assess_ramu_fits <- read_rds(file = file.path(results_path,"assess_ramu_fits.rds"))

ram_v_sraplus_area <- read_rds(file = file.path(results_path,"ram_v_sraplus_area.rds"))

ram_comp_data <- read_rds(file = file.path(results_path,"ram_comp_data.rds"))

fao_sraplus_comp_data <- read_rds(file = file.path(results_path, "fao_sraplus_comp_data.rds"))

load(file = file.path(results_path, "paper_plots.Rdata"))

load(file = file.path(results_path, "voi_fits.RData"))

  exs <- read_rds(file = file.path(results_path, "raw-case-study-fits.rds"))

breaks <- c(0.8, 1.2)

breaks <- c(0, breaks, Inf)

labels <- c("over", "fully", "under")

# rand_fits <- assess_ram_fits %>%
#   filter(data == "ram-data") %>%
#   mutate(mean = sample(c(0.1, 1, 2), n(), replace = TRUE),
#          data = "guess")

perf <- assess_ram_fits %>% 
  # bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_b_v_bmsy, breaks = breaks, labels = labels)) %>%
  mutate(data = as.factor(data)) %>% 
  filter(data != "u_umsy") %>% 
  mutate(data = forcats::fct_recode(data, 
                                    "RLSADB Index" = "ram-data",
                                    "SAR" = "sar",
                                    "FMI" = "fmi",
                                    "Effective CPUE" = "cpue",
                                    "Effective CPUE+" = "cpue-plus",
                                    "Nominal CPUE" = "nominal-cpue",
                                    "Nominal CPUE+" ="nominal-cpue-plus",
                                    "CMSY" = "cmsy",
                                    "Guess" = "guess",
                                    "RLSADB F/Fmsy" = "u_umsy"

                                    )) %>% 
  group_by(data) %>% 
  summarise(
    mpe = median((mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    mape = median(abs(mean - ram_b_v_bmsy ) / ram_b_v_bmsy),
    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
  ) %>% 
  arrange((mape)) %>% 
  rename("Data Used" = "data")

uperf <- assess_ramu_fits %>% 
  # bind_rows(rand_fits) %>% 
  mutate(sraplus_bin = cut(mean, breaks = breaks, labels = labels)) %>%
  mutate(ram_bin = cut(ram_u_v_umsy, breaks = breaks, labels = labels)) %>%
  mutate(data = as.factor(data)) %>% 
  filter(data != "u_umsy") %>% 
  mutate(data = forcats::fct_recode(data, 
                                    "RLSADB Index" = "ram-data",
                                    "SAR" = "sar",
                                    "FMI" = "fmi",
                                    "Effective CPUE" = "cpue",
                                    "Effective CPUE+" = "cpue-plus",
                                    "Nominal CPUE" = "nominal-cpue",
                                    "Nominal CPUE+" ="nominal-cpue-plus",
                                    "CMSY" = "cmsy",
                                    "Guess" = "guess",
                                    "RLSADB F/Fmsy" = "u_umsy"

                                    )) %>% 
  group_by(data) %>% 
  summarise(
    mpe = median((mean - ram_u_v_umsy ) / ram_u_v_umsy),
    mape = median(abs(mean - ram_u_v_umsy ) / ram_u_v_umsy),
    accuracy = mean(ram_bin == sraplus_bin, na.rm = TRUE)
  ) %>% 
  arrange((mape)) %>% 
  rename("Data Used" = "data")

fao_area_codes <- fao %>% 
  ungroup() %>% 
  select(fao_area, fao_area_code) %>% 
  unique()

pub_theme <- theme_ipsum(base_size = 10,
                         axis_text_size = 10,
                         axis_title_size = 12) +
    theme(panel.spacing = unit(.5,"lines"))

theme_set(pub_theme)

```

# Title Page {.unnumbered}

**Title 1**: Improving Estimates of the State of Global Fisheries Depends on Better Data

**Title 2**: Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

**Running Title**: Unassessed Fisheries

**Authors**:
Daniel Ovando^1^, Ray Hilborn^1^, Cole Monnahan^2^, Merrill Rudd^3^, Rishi Sharma^4^, James T. Thorson^2^, Yannick Rousseau^5^, Yimin Ye^4^

**Affiliations**:

^1^University of Washington, School of Aquatic and Fishery Sciences
1122 NE Boat St, Box 355020 Seattle, WA 98195-5020
Seattle, WA, USA 98103

^2^NOAA Fisheries Alaska Fisheries Science Center, Seattle, WA, USA

^3^Scaleability LLC

^4^Food and Agriculture Organization of the United Nations, Fisheries and Aquaculture,Rome, IT

^5^University of Tasmania, Institute for Marine and Antarctic Studies,Hobart, TAS, AUS

**Authorship**: DO, RH, CM,MR, RS, JT designed model structure and ran analyses. RS, YR, and YE supplied data. All authors contributed to writing of the manuscript

**Conflict of Interest**: RH receives research funding from many groups that have interests in fisheries outcomes including environmental NGOs, foundations, governments and fishing industry groups.

\newpage

# Abstract {.unnumbered}

Implementation of the United Nations Sustainable Development Goals requires assessments of the global state of fish populations. While we have reliable estimates of stock status for fish populations accounting for approximately half of recent global catch, our knowledge of the state of the majority of the world's 'unassessed' fish stocks remains highly uncertain. Numerous publications have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. Here, we show that despite numerous efforts, our understanding of the status of global fish stocks remains incomplete, even when new sources of broadly available data are added. Estimates of fish populations based primarily on catch histories on average performed 25% better than a random guess. But, on average these methods assigned fisheries to the wrong FAO status category 57% of the time. Withing these broad summaries the performance of models trained on our tested data sources varied widely across regions. Effective improvement in estimates of the state of the world's exploited fish populations depends more on expanded collection of new information and efficient use of existing data than development of new modeling methods.

# Introduction {.unnumbered}

The United Nations Sustainable Development Goal 14 (SDG 14), focusing on "Life under water", calls for the global community to "Conserve and sustainably use the oceans, seas and marine resources for sustainable development". Marine fisheries are one of the largest anthropogenic impacts in the oceans aside from the effects of climate change. Fisheries are also a critical source of economic prosperity, cultural identity, and food security around the globe. As such, meeting the SDG 14 targets depends in part on our ability to effectively measure the status of global marine fish populations and fisheries. While our understanding of the state of world fisheries has improved over the last decade, the majority of the world's fish populations, making up roughly 50% of marine landings (though only a few percent of the total number of fisheries in the world), lack formal statistical assessments (stock assessments) of their population size relative to reference points [@hilborn2020; @fao2020]. This is a major impediment to ensuring the sustainable development of the world's oceans. In this paper, we consider why the assessment of global fisheries remains a challenge, and chart a path towards a better understanding of the state of the world's marine resources.

What do we currently know about the state of fished marine species around the world? The Food and Agriculture Organization of the United Nations' (FAO) is the custodian agency of the SDG 14 Indicator on fisheries sustainability, and the FAO's State of World Fisheries and Aquaculture (SOFIA) report is the most widely used primary source for tracking the global state of fisheries. The latest SOFIA report, covering 70% of the landings of all fisheries in the world, estimates that as of 2017 59.6% of marine fish stocks are maximally sustainably fished (at or near targets for sustainable food production), 6.2% are underfished, and 34.2% are overfished [@fao2020]. Where possible, these statements about the status of individual fish stocks are made on the basis of formal stock assessments summarized in the RAM Legacy Stock Assessment Database (RLSADB) [@ricard2012]. The "assessed" fisheries in RLSADB represent roughly 50% of global capture as of 2020 [@hilborn2020], and represent our best estimates of the state of assessed fish populations around the globe.

That leaves roughly 50% of global landings, and the majority of global fisheries, as currently "unassessed". While these unassessed stocks are generally individually smaller and less economically valuable than the fish populations in the assessed category, collectively they are a vital source of food, employment, cultural value, and ecosystem services around the world. The SOFIA report includes unassessed stocks and bases its estimates for these fisheries on data-limited methods or qualitative expert opinion for each region where these stocks are distributed [@fao2020]. While these methods combined with local knowledge can provide good insight as to the general status of unassessed fish stocks, the SOFIA assessment was designed in the 1970s based on the then available data and methods. With the surge in data availability, such as global assessments of management strength [@melnychuk2017], trawl footprints [@amoroso2018], and fishing effort [@rousseau2019], the SOFIA global assessment methods for unassessed fish populations need to be updated to meet the demand for tracking progress of global fish populations towards the SDG goals. In addition, the SOFIA process assigns unassessed fisheries to broad categories of stock status, while many stakeholders seek information on more specific values such as biomass relative to the biomass that would on average result in maximized yield over the long-term.

<!-- We evaluate the ability of alternative assessment models to improve our estimate of the population status of unassessed fish stocks around the world.  -->

Numerous studies in recent years have put forward versions of "data-limited" models that have attempted to provide numerical estimates of the global status of unassessed fish stocks lacking the data or capacity needed for formal stock assessment [@costello2012b; @costello2016a; @pauly2007; @rosenberg2018; @thorson2012a]. Due to data limitations, all of these global assessment efforts have used forms of "catch-only" data-limited models (@free2020 and references therein). These models seek to infer the state of a fished population, for example biomass (*B*) relative to the biomass at maximum sustainable yield (*B~MSY~*, the ratio *B/B~MSY~* being a common measure of stock status), from characteristics of a fishery's catch history.

<!-- These issues become apparent when we consider some of the macro-level predictions made by these models. While the estimates of *B/B~MSY~* and other reference points reported in RLSADB are themselves model outputs subject to their own non-trivial errors and biases, a simple benchmark is to compare the estimates of fishery status from RLSADB to those predicted by catch-only models intended for use when insufficient data are available for a full stock assessment model. -->

<!-- @costello2016a found similar rankings of regions in terms of stock status as RLSADB, but their estimate of the state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seems to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in @rosenberg2018 demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. Simply comparing catch divided by maximum catch, in the manner of @pauly2007, finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RLSADB results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly exploited regions as measured by RLSADB and the FAO resulting from these catch-only efforts at assessing global fisheries is concerning. -->

```{r status-tab, include=FALSE, eval=FALSE}

o_by_a <- sofia_status %>% 
  left_join(fao_area_codes, by = "fao_area_code") %>% 
  group_by(fao_area) %>% 
  summarise(`FAO % Overfished` = percent(mean(status == "O")))

ss <-
  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
                    sheet = "state_space_results_08_28_2018") %>%
  janitor::clean_names()

ss_lookup <-
  readxl::read_xlsx(here("data", "state_space_results_08_28_2018 by region.xlsx"),
                    sheet = "ss_to_fao") %>%
  janitor::clean_names() %>% 
  mutate(tmp = map2(region, fao_area_codes, ~data.frame(region = .x, fao_area_code = as.integer(str_split(.y, pattern = ',',simplify = TRUE))))) %>% 
  select(tmp) %>% 
  unnest(cols = tmp) %>% 
    left_join(fao_area_codes, by = "fao_area_code")

state_space_results <- ss_lookup %>% 
  left_join(ss, by = "region") %>% 
  filter(variable == "BvB") %>% 
  group_by(fao_area) %>% 
  filter(year == max(year)) %>% 
  summarise(`B/BMSY RLSADB` = round(mean(dlm_geomean),2)) %>% 
  ungroup()

# ram_by_region %>% 
#   ggplot(aes(b_v_bmsy)) +
#   geom_histogram() + 
#   facet_wrap(~fao_area)

tmp <- readr::read_csv(here("data","status-table.csv")) %>% 
  select(-`B/BMSY RLSADB`) %>%  
  left_join(o_by_a, by = c("FAO Area" = "fao_area")) %>% 
  left_join(state_space_results, by = c("FAO Area" = "fao_area")) %>% 
  arrange(`FAO % Overfished`) %>% 
    rename(`SOFIA % Overfished` = `FAO % Overfished`) %>% 
  select(`FAO Area`,`SOFIA % Overfished`,`B/BMSY RLSADB`, everything())

 tmp %>% 
  knitr::kable(caption = "Estimates of fish stock status by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, Catch / max(Catch)",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

```

However, @free2020 demonstrated that the types of catch-only models used in these global assessment efforts can often produce both imprecise and biased estimates of current stock status in terms of *B/B~MSY~*. While each of these prior efforts at assessing global fisheries using catch-only methods made important advances in our understanding of the global oceans, none have proven to be a consistently reliable means of estimating the state of the individual unassessed fisheries around the world. The types of global assessment efforts share a common underlying assumption; that globally available catch histories contain meaningful information on the state of the fished populations that they came from, and that this information can be revealed by the right model structure. Under this assumption, we can obtain better estimates of the state of unassessed fish populations simply by improving the catch-only models applied to the catch histories from these fisheries. However, further evaluation of the predictions made by these models [@free2020;@bouch2021;@pons2020] as well as works such as @pauly2013 and @branch2011 call into question this very idea.

In this paper we ask, how much can estimates from catch-only assessments be improved by moving beyond catch-only and augmenting catch histories with additional sources of broadly available fisheries data? We answer this question by aggregating a set of broadly distributed datasets that might be of use to global fishery assessment. We then used our new assessment method, sraplus, to evaluate the performance status estimates derived from different combinations of these broadly available data. We compared these results with estimates from the increasingly utilized (e.g. @palomares2020) default settings of the catch-only version of CMSY described in @froese2017. We show that improving our understanding of the world's unassessed fisheries depends on a redoubled effort at global data collection and synthesis, not on incremental improvements to assessment methods based solely on catch histories.

# Methods {.unnumbered}

The basis of this analysis is a new stock assessment software package we call [sraplus](https://github.com/DanOvando/sraplus) (https://github.com/DanOvando/sraplus). sraplus is an extension of stochastic stock reduction analysis (SRA) [@kimura1984;@walters2006], which allows users to combine a biomass dynamics model with a variety of data sources (e.g. priors on recent stock status, or an index of abundance) in order to produce estimates of the state of a fishery over time. The key goal of sraplus is not substantial improvements in model fitting methods per say, but providing the ability to easily incorporate multiple kinds of fishery data potentially used in SRA-style analyses in a statistically rigorous manner. We paired various types of data with fisheries in RLSADB, and then used sraplus to generate predictions of *B/B~MSY~* in the most recent year of each fishery across different combinations of data types. We then compared the predictions generated by sraplus using different types of data, along with predictions for the same fisheries generated by CMSY, to the reported values in RLSADB.

All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries). Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.

## Data Sources {.unnumbered}

At its most "data-limited", sraplus can work as a catch-only model in the manner of @froese2017. These catch histories can then be augmented with prior information on stock status, derived from expert opinion, or using built-in prior generating models based on increasingly available sources of global fishery data: the Fisheries Management Index (FMI) [@melnychuk2017] and swept area ratio (SAR) [@amoroso2018] databases. The FMI database provides self-reported scores reflecting the strength of fisheries management in many fisheries around the world, which sraplus uses to construct informative priors on stock status, e.g. placing a higher probability that fisheries with high FMI scores have better stock status than those with low FMI scores. SAR is a measure of the intensity of trawl fishing within a particular area, which sraplus uses to construct an informative prior on the magnitude of fishing mortality. While fishery-independent surveys are becoming increasingly available [@maureaud2020], they are not yet sufficiently distributed or accessible to serve as a foundation for global assessments of unassessed fisheries. As an alternative, @rousseau2019 present a global reconstruction of fishing effort for countries around the world. We pair these effort reconstructions (both nominal effort and effective effort assuming an annual rate of efficiency increase of 2.6%) with the FAO's catch histories to construct a catch-per-unit-effort (CPUE) index, which we pass to sraplus along with catch histories and other available data to provide an estimate of stock status. In this way sraplus allows us to move from catch-only models to more conventional surplus production models in the manner of @winker2018 (Table.\@ref(tab:data)).

<!-- In this paper we ask, can augmenting the types of catch-only models employed in global assessments to date with new broadly available data improve our understanding of the state of global fish populations? We use a flexible stock assessment software package, [sraplus](www.github.com/danovando/sraplus) to demonstrate how different data sources can be used to augment catch-only models at a global scale, and to evaluate how our perception of global stock status would vary depending on which sources of data we include. -->

<!-- ## Experiment Structure {.unnumbered} -->

<!-- We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database.  -->

<!-- ## Data Sources {.unnumbered} -->

<!-- We restricted ourselves to databases with wide coverage around the world, representing the types of information that can be leveraged in efforts to assess the status of global fisheries (Table.\@ref(tab:data)).  -->

## Population Model {.unnumbered}

The core of sraplus is a Pella-Tomlinson [@pella1969] production model constructed in the manner of @winker2018. While models of these kinds abstract away many important details of fish biology and fleet behavior, they are the highest resolution model that the data types evaluated here will support.

<!-- The purpose of sraplus is not to make substantial innovations in the fitting of surplus production models, but to provide a flexible tool for exploring the impacts of adding different kinds of data and priors on estimates of fishery status -->

The population growth equation is

\begin{equation}
f(x)=\begin{cases}
B_{t + 1} = \left(B_{t} + B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)p_t
, & \text{if $B_t>0.25 \times K$}.\\
B_{t + 1} = \left(B_{t} + \frac{B_{t}}{0.25 \times K}\left(B_{t}\frac{r}{m - 1}\left(1 - \left(\frac{B_t}{K}\right)^{m- 1}\right) - c_t\right)\right)p_t, & \text{otherwise}.
\end{cases}
(\#eq:pt)
\end{equation}

Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift. When *m* = 2, *B~MSY~ / K* = 0.5. Lower values of *m* shift the production function left, higher values right. The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, @thorson2012 provides estimates of the ratio of *B~MSY~* to *K* for many fish taxa. For each stock we fix the shape parameter based on the distributions reported in @thorson2012 for the genus of the species in question. We chose to fix the shape parameter at the mean stock-appropriate values from @thorson2012 rather than estimating the shape parameter with an informative prior since there is so little information regarding the shape in the data considered. Attempts to estimate the shape parameter with priors from @thorson2012 frequently resulted in poor model performance. $\pmb{c}$ is a vector of catches, and $\pmb{p}$ is vector of process errors. Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model. We dealt with this problem by following the methods described in @winker2018 to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.

We allow for process error $p_t$ (in the manner of the stochastic stock reduction analysis (SRA) suggested by @walters2006). Process error $p_t$ is assumed to be log-normally distributed, such that

<!-- This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]). -->

\begin{equation}
  log(p_t) \sim {N}\left(-\sigma_{proc}^2/2,\sigma_{proc}\right)
(\#eq:procerror)
\end{equation}

where *N* is the normal distribution.

## Estimation Model {.unnumbered}

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011]. Unless there is an abundance index to fit to, the model runs as an SRA. A stock reduction analysis works by specifying prior distributions on population parameters and, critically, the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. All stock reduction analysis runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e^6^ candidate draws. For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r* (with a prior distribution drawn from Fishlife, @thorson2020), *K*, $\sigma_{proc}$, and *B~0~* (initial depletion, $B/K$). *q* is also estimated when needed. $\sigma_{proc}$ (process error) is estimated indirectly through the parameter $\gamma$, the ratio of process (*proc*) to observation (*obs*) error ($\sigma_{obs}$). See Table.S1 for prior distributions for each of these parameters.

<!-- . We can break the use of sraplus into two distinct categories: with data to fit to and without. By "data", we refer to measurements which are used to confront model estimates within a likelihood function. In our context, these include fishery-independent survey data, or a CPUE index. When there are no data to fit to, the model amounts to filtering priors through the model (the combination of the Pella-Tomlinson operating model and the catches for the stock in question, along with any fixed parameters). Under this mode, the model is essentially a stock-reduction analysis model, in the manner of [@walters2006], in which we ask, which combinations of prior probability distributions of parameters do not crash the population (i.e. results in biomass less than catches in any time step in the fishery's history), given the constraints of the population model and the catches. This step updates the prior distribution of population parameters by eliminating combinations of priors that are impossible for a given catch history and a specified functional form. -->

<!-- The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).  -->

<!-- *r* and *K* are the only two parameters that are always estimated. Estimation of every other parameter can be turned on or off. When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user. -->

<!-- sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent). Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results based on data sources listed in Table.\@ref(tab:dat-desc) that do not contain "cpue" or "RAM data" are estimated through stock reduction analysis. All SRA style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws. -->

When an index of abundance is available the model estimates the posterior probability distributions of the estimated and transformed parameters using Hamiltonian Monte Carlo implemented in Stan [@standevelopmentteam2018] accessed through the `tmbstan` interface [@monnahan2018]. By default the model uses 2000 draws with a 1000 step warm-up and one chain. Any detailed fit for an individual fishery would likely use more draws and chains, but we verified that this sampling routine produced an acceptable tradeoff of speed and convergence criteria. The model fits to a direct estimate of abundance (e.g. a fishery independent survey or a standardized catch-per-unit-effort index), the likelihood calculated per

\begin{equation}
  B_t = f_{t}(r,K,m,B_0,\pmb{p},\pmb{c})
(\#eq:biomass)
\end{equation}

where $B_t$ is the biomass in time *t*, given $f_{t}$ which is shorthand for the result of the Pella-Tomlinson equation in time step *t* given parameters and a vector of catches. The likelihood of the abundance index *a* in time *t* is then calculated as

\begin{equation}
 log(a_t) \sim N(B_t \times q, \sigma_{obs})
(\#eq:likelihood)
\end{equation}

When an effort index is available, sraplus constructs an index of abundance based on the catch and effort data. @rousseau2019 measure an index of abundance as catch divided by their effort index, either nominal or effective (assuming the 2.6% annualized technology creep). Treating this raw effort data as the denominator in the CPUE calculation assumes that every increase in fishing effort translates to a commensurate increase in fishing mortality. When effort increases dramatically above historic levels, this can create a CPUE index that decreases faster than the true population. This is due to the fact that in reality the marginal fishing mortality produced by increasing units of effort decreases as effort approaches infinity (since the realized exploitation fraction, sometimes called *u*, is between 0 and 1). To accommodate this, we generate a catch per effective harvest rate index of abundance, as

\begin{equation}
 cpue_t = \frac{c_t}{(1 - e^{-F_t})}
(\#eq:cpue)
\end{equation}

\begin{equation}
 F_t = q_tE_t
(\#eq:f)
\end{equation}

Where $q_t$ can has a technology creep component $\tau$

\begin{equation}
 q_t = q_{t-1} \times (1  + \tau)
(\#eq:qt)
\end{equation}

We then fit to the index of abundance per Equation 3 and

\begin{equation}
 log(cpue_t) \sim N(B_t, \sigma_{obs})
(\#eq:cpue)
\end{equation}

### CMSY {.unnumbered}

In addition to the results from sraplus, we include a set of results produced by the default settings of the CMSY method [@froese2017]. For computational efficiency, we used a ported version of the CMSY model available at <https://github.com/DanOvando/portedcmsy>. The only modification made is to convert the underlying population model to C++ for faster computation. For each stock we used all the default options and priors provided and generated by CMSY, in the same manner as @palomares2020, except for resilience, which was pulled from the vulnerability scores from FishBase accessed through `rfishbase` [@boettiger2012]. Vulnerability scores greater than 66 were scored as low resilience, between 33 and 66 medium resilience, and lower than 33 high resilience.

<!-- In the absence of an index of abundance CMSY and sraplus are conceptually the same: Both sample from prior distributions of life history and stock status conditional on the stock not collapsing given a catch history and a functional form for the population growth equation. The core different between CMSY and sraplus as utilized here then is that when we refer to CMSY we mean CMSY run using the default heuristics used by the algorithm to construct priors on initial, intermediate, and terminal stock status. In contrast, sraplus fit using for example FMI data uses the empirical models described here to generate priors on stock status. In addition, all sraplus runs use a prior-predictive tuning procedure to resolve issues related to Borel's paradox, described below. -->

## Priors {.unnumbered}

<!-- We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus. -->

### Prior Predictive Tuning {.unnumbered}

<!-- We can think of this as a binomial process in which we fit a model of the population, conditional on catches, to the fact that we know that population existed in each time step of the catch history. Beyond that though, baring additional information or model assumptions we have no way of knowing whether these catches represent a substantial proportion of a small population or a minuscule fraction of a massive population; all we know is that the population must always have been above zero. -->

<!-- Suppose that the only thing we observe from a fishery is a catch history. Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.  -->

In the absence of any data to fit to, sraplus works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history, life history priors, and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See @poole2000 and references therein for a discussion of Borel's Paradox in a fisheries context).

This may seem like an academic concern, and indeed in our experience when the data are sufficiently informative the Bayesian version of our model subject to Borel's paradox produces effectively identical results to those produce by the same model fit by maximum likelihood. However, Borel's Paradox poses a particular problem when there are no data to fit to(i.e. when the model is simply filtering through prior distributions in the manner of a traditional SRA) due to the fact that there are more parameter combinations that allow for a fishery to be relatively unexploited than for a fishery to be close to collapse (but never actually collapsed, i.e. predicted biomass less than observed catch). In this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although combined with other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the posterior distributions of stock status resulting from the SRA algorithm.

We use an approximate solution to this problem here, similar in spirit to Bayesian melding [@poole2000]. Our solution amounts to a two-step sampling-importance-resampling (SIR) algorithm. We first run the standard SRA algorithm as described in the Estimation Model section of the methods. We then break the resulting draws into bins based on terminal stock status, and calculate the mean probability density *p* (defined by the prior distributions of estimated parameters) of each bin.

\begin{equation}
 p(bin_i) = \frac{1}{N_i}\sum_{n = 1}^{N_i}{p(b_{n,i})}
(\#eq:pbin)
\end{equation}

We then divide the mean probability density of bin *i* evenly among each of the draws within that bin *n*

\begin{equation}
 p(n_i) = \frac{p(bin_i)}{N_i}
(\#eq:ppcor)
\end{equation}

And we then perform a second SIR algorithm but now sampling each observation $n_i$ in proportion to $p(n_i)$.

The net result of this is that it allows users to place an explicit prior on stock status, and then adjust their priors on life history parameters to reflect this prior. While the range of possible life history values supplied still influences stock status under this approach, this prior predictive tuning process makes the resulting priors more consistent with explicit priors on recent stock status supplied by the user. Users can turn this functionality off and instead base priors on stock status primarily on life history. See Supplementary Information for a detailed explanation of this problem and our solution.

### Priors Informed by Outside Data {.unnumbered}

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio, and Fisheries Management Index with estimates of stock status from the RLSADB. We then trained a regression of the general form $log(status) \sim N(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See Supplementary Information). All prior regression models were tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus. For all results presented in this paper we used these data to provide priors on *F/F~MSY~*, as we found clearer predictive relationships and subsequent model performance between catch, FMI, and SAR values and *F/F~MSY~* than we did for *B/B~MSY~*.

<!-- Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here. -->

## Assessing Performance {.unnumbered}

Simulation testing can be preferable in many ways to comparison against model outputs. However it is not possible to simulation test the value of information derived from empirical relationships between variables such as fisheries management strength and fishery outcomes, which our study depends on. As such we assess model performance through comparison to best available estimates of stock status available in RLSADB.  We based this test on `r ram_comp_data$stockid %>% n_distinct()` stocks from RLSADB, covering `r n_distinct(ram_comp_data$isscaap_group)` broad taxonomic groups, with estimates of *B/B~MSY~* and greater than 25 years of continuous catch history. *B/B~MSY~* values from RLSADB are themselves estimates, not data, but they are the best available information on global stock status. We then paired the catch histories for these RLSADB stocks with regional-level SAR, FMI, and effort data.  Our methods approximated a regional-level assessment exercise, where data beyond catch histories are available at regional levels, but not for specific fisheries. We also estimated *B/B~MSY~* values of our candidate RLSADB stocks by using sraplus to fit to an abundance index drawn directly from RLSADB itself "RLSADB Index", as a measure of the ability of models like sraplus if given perfect information

We then fit a range of models utilizing different combinations FMI, SAR, and effort data, along with CMSY, and a set of runs fit to the RLSADB Index (Table.\@ref(tab:dat-desc)). We assessed model performance using three metrics: median percent error (MPE, a measure of bias), median absolute percent error (MAPE, a measure of accuracy), and classification accuracy. Classification accuracy was calculated as the proportion of times that use of a given combination of data resulted in a stock being classified into the correct FAO status classification (one of underfished, maximally sustainably fished, and overfished). This experiment structure allows us to assess the ability of catch-only models to estimate the *B/B~MSY~* values reported in RLSADB, and evaluate how much these estimates can be improved by augmenting the catch histories with additional sources of information. Our benchmark model is a simple "Guess". Guess assigns each stock a random *B/B~MSY~* of 1.6, 1, or 0.4, corresponding to the middle value of the FAO status bins of underfished  ($B/B~MSY~ \geq 1.2$), maximally sustainably fished ($0.8 \geq B/B~MSY~ < 1.2$), overfished  ($0 \geq B/B~MSY~ < 0.8$). We performed a matching analysis measuring performance in estimating *F/F~MSY~* as well, with results presented in the SI.

## Value of Information Calculations {.unnumbered}

We performed a value of information assessment to determine what types of data may be most beneficial to acquire at a global scale if we are to improve our knowledge of the state of global fisheries. The value of information analysis was performed by using sraplus to generate estimates of *B/B~MSY~* for stocks in RLSADB, and comparing the estimated values to the values reported in RLSADB. There are too many combinations for us to run the full expansion of possible parameter states. To resolve this we generated fits for 3000 combinations of a RLSADB stock and available data. For any one draw, we randomly sampled a RLSADB stock and a list of available data and data quality. For example, we might sample stock *A* with information on recent fishing mortality rates for the first iteration, and stock *A* again for the second iteration but now with information on recent fishing mortality rates and a recent index of abundance. The result is a set of model performance estimates where the characteristics of the stock and the data made available to the model are randomized. We then measured the value of information as the average reduction in root mean squared error (RMSE) in *B/B~MSY~* over the most recent five years of the fishery (in order to evaluate the ability of the model to capture recent trends as well as the most recent value), resulting from use of different kinds of data. We considered the value of information of having data on: the most recent *B/B~MSY~*, treating initial state of the population as one of unfished ($B_{0}=1$), known ($B_{0}$ taken from stock assessment), or estimated based on initial shape of the catch history, *F/F~MSY~* values in the most recent year, over the last five years, and over the complete time series, and an abundance index spanning the complete, most recent half, or most recent quarter of the time series. We considered the value of information of a longer time series of *F/F~MSY~* than *B/B~MSY~* to consider the potential of augmenting catch-only models with additional data that can inform *F/F~MSY~*, such as length-composition data, that might be available over the history of the fishery.

## Case Study {.unnumbered}

We also ran a case study demonstrating how different kinds of data led to different conclusions about stock status. We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from @rousseau2019 to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the default settings of the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used sraplus to generate estimate of stock status based solely on catch heuristics, taking into account the prior-predictive tuning implemented by default in sraplus. We also used stock-specific data on SAR and FMI to generate priors on *F/F~MSY~* for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. This case study represents a more "localized" assessment, where external data sources (FMI and SAR) are available at the stock level, rather than at the regional level. We summarized the performance of each model fit in the case study based on root mean squared error (RMSE).

<!-- To determine what sources of data might provide the greatest value in improving our estimates of global stock status, we fit 3,000 sraplus models to randomly sampled fisheries from RLSADB, each time varying the kind and quality of data made available to the model, and what parameters the model estimated. We then calculated the root-mean-squared-error (RMSE) between the observed and predicted B/B~MSY~ over the most recent five years of each fit, and then fit a regression to these data to estimate the posterior probability of the effect of different data types and model states on RMSE (Fig.\@ref(fig:voi-plot)). -->

# Results {.unnumbered}

## Case Study {.unnumbered}

Nearly all of the fisheries used in this case study have *F/F~MSY~* values less than one, and most have *B/B~MSY~* values greater than one. Both the catch heuristic implemented in sraplus and CMSY, using no stock-specific information but the catch history (which defines the prior on K and stock depletion, and hence the estimate of *B/B~MSY~*) and the estimated resilience of the target species used to provide a prior on *r* (based on Fishlife, @thorson2020), badly miss this trend, predicting instead that nearly all the fisheries in this group are currently overfished and experiencing overfishing (Fig.\@ref(fig:cs-plot)). Overall, models informed by the SAR and FMI values performed the best in terms of *B/B~MSY~* (RMSE = 0.89). The ranking of performance by data sources used was the same for *F/F~MSY~*, though the range of values was much higher, with the SAR and FMI based result producing an RMSE of 0.39 while the catch-only case studies produced *F/F~MSY~* RMSE values greater than 1 (Fig.\@ref(fig:cs-plot)). However, this improved performance of the SAR and FMI data is to be expected as the prior-generating models were trained on data from RLSADB, and therefore this is likely an optimistic assessment of the performance of these prior-generating models when applied to new fisheries.  Fitting to an index of abundance created by regional effort data in fact produces a slightly poorer fit in *B/B~MSY~* and *F/F~MSY~*. This is likely because a number of stocks in this case study set have very low catches (relative to historic highs) but high biomass values, creating a mismatch between the CPUE trends based on the effort and catch data and actual biomass. However none of the models fit to the four case-study data combinations were able to explain much of the variation in *B/B~MSY~*, though the models performed better in some cases for predicting *F/F~MSY~*.

<!-- This is not in and of itself a surprising result, but demonstrates how 1) estimates of stock status based on catch histories alone can be both imprecise and biased and how 2) augmenting catch histories with additional data from the fisheries in question can produce meaningful improvements.  -->

<!-- `FishLife` [@thorson2017c] provides informative priors for key life history parameters (such as growth rates) needed by our model. We use the FMI and SAR values to construct Bayesian priors on initial and terminal stock status values (e.g B/B~MSY~ or F/F~MSY~, the fishing mortality *F* relative to the fishing mortality rate that would produce *MSY* at equilibrium), allowing us to translate local values of for example SAR into a prior on terminal B/B~MSY~. Where available, we use effort data to create stock-specific timeseries of catch-per-unit-effort (CPUE), which can be used as an index of stock abundance.  -->

```{r cs-plot, fig.cap="RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], catch heuristics, priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort (Effort) trends assuming a rate of technological increase of 2.6%. Each point is a stock in the RLSADB. Black dashed line shows the 1:1 relationship. Text displys root mean squared error (RMSE) of each panel."}
ex_scatter_plot + 
  theme(strip.text = element_text(size = 8))

ggsave(here("documents","figs","cs-plot.png"),ex_scatter_plot, height = 6, width = 6)

```

<!-- That use of Swept Area Ratio and Fisheries Management Index data produced a more accurate estimate of stock status than catch alone should come as no surprise. In the absence of other data, the heuristics internal to CMSY linking catch to stock status, and the regressions linking SAR and FMI values to stock status, are themselves the assessment process. Since the SAR and FMI regressions were trained on the RLSADB stocks in question, this process simply confirms that for these stocks SAR and FMI data contain a reasonable amount of information on fishing pressure and by extension stock status. -->

## Performance of Regional Fishery Assessments {.unnumbered}

We first considered the performance of sraplus when the model was given a perfect index of abundance from RLSADB (RLSADB Index) for each stock in RLSADB it was tasked with assessing. Assessment models in RLSADB are typically much more structurally complex than the simple biomass dynamics used in sraplus, and so this "perfect information" test tells us how much of a penalty in model performance we are likely to pay due to model misspecification alone. The sraplus estimates of *B/B~MSY~* resulting from fitting directly to the abundance indices from RLSADB were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "RLSADB Index"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "RLSADB Index"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "RLSADB Index"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality index of abundance, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

<!-- Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves.  -->

 <!-- However, this is somewhat an artifact of the data. The status of most stocks in RLSADB is relatively good, with recent *B/B~MSY~* values generally near one. This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise. -->

We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status beyond those derived from catch-only methods. Many of the datasets used produced similar levels of bias as the RLSADB data (e.g. FMI, SAR, Nominal CPUE fits), though notable the "Guess" method actually performed near the top in terms of bias. However, this is somewhat an artifact of the data. The status of most stocks in RLSADB is relatively good, with recent *B/B~MSY~* values generally near one. As the mean value of the "Guess" method is one, on average the "Guess" model is an unbiased but imprecise measure of stock status in RLSADB (Table.\@ref(tab:perf-tab)).

Focusing on MAPE (our measure of error rather than bias) and classification accuracy, the error of the models jumps dramatically as soon as data other than the RLSADB abundance indices are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy of the sraplus models across all non-RLSADB data fits was `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))`. Note that there are only three bins in the FAO stock status classifications, and as such our "Guess" model has a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "Guess"])`. This means that the accuracy of our models designed as a proxy for a global assessment process were across all non-RLSADB index data fits `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]) / perf$accuracy[perf$"Data Used" == "Guess"] - 1)` more accurate than a random guess, certainly an improvement, but on average assigned fisheries to the wrong FAO status bin `r percent(1 - mean(perf$accuracy[!perf$"Data Used" %in% c("RLSADB Index","Guess")]))` of the time.

Looking geographically we found a similar pattern of a rapid decrease in performance for models besides those fit to the RLSADB Index. Across the models, performance was not consistent in space: use of different data performed best or worst for different FAO regions. Models fit to nominal CPUE data substantially overestimate stock status in the Mediterranean, while models based on data using effective CPUE perform better in that region (but worse in others) (Fig.\@ref(fig:mape-map)). We find similarly inconsistent performance for both bias (Fig.\@ref(fig:mpe-map)) and accuracy (Fig.\@ref(fig:acc-map)). Overall, while some data sources performed slightly better than others by some metrics in some places, no models using any non-RLSADB index data were able to capture the overall state or geographic distribution of stock status represented in RLSADB in a consistent manner. Performance in estimating *F/F~MSY~ * was similarly variable and poor, with the exception that the default settings of CMSY performed much more consistently poorly in terms of *F/F~MSY~* than *B/B~MSY~* (due to systemic overestimation of *F/F~MSY~*, see Supplementary Information [SI]).

<!-- xx dig into this, which stocks are in that region? XX -->

```{r mpe-map, fig.width=8, fig.height=6, fig.cap="Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources. Data source panels are ordered in ascending total bias at the FAO region level. RLSADB Index refers to catch and abundance index drawn from RLSADB. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6. "}
ram_mpe_map_plot

ggsave(here("documents","figs","mpe-map.png"), ram_mpe_map_plot, width = 8, height = 6)
```

```{r mape-map, fig.cap = "Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources. Data source panels are ordered in ascending (starting from top left) mean MAPE at the FAO region level. RLSADB Index refers to catch and abundance index drawn from RLSADB. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}
ram_mape_map_plot

ggsave(here("documents","figs","mape-map.png"),ram_mape_map_plot, width = 8, height = 6)

```

```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources. Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level. RLSADB Index refers to catch and abundance index drawn from RLSADB. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.",fig.width=8, fig.height=6}

ram_acc_map_plot

ggsave(here("documents","figs","acc-map.png"),ram_acc_map_plot, width = 8, height = 6)

```

\newpage

<!-- The poor global performance of models fit to currently accessible broadly available datasets, relative to models fit to data from RLSADB, indicates that we must prioritize collection and aggregation of new or underutilized data sources if we are to improve our estimates of global fishery status.  -->

## Value of Information {.unnumbered}

Having access to estimates of *F/F~MSY~* reduced model error in proportion to the number of years for which *F/F~MSY~* values are available. Interestingly though, having access to only an accurate estimate of *F/F~MSY~* in the most recent year was extremely informative, reducing error on average by 15%, on par with an estimate of recent *B/B~MSY~* itself  (Fig.\@ref(fig:voi-plot)). While having access to complete index of abundance, such as a fishery independent survey, was on average able to reduce error relative to a baseline catch-only heuristic, using only the most recent quarter of the available abundance index actually increased error on average, due to the lack of historical context for recent trends in abundance.

<!-- We may have to wait many years for new surveys to provide substantial improvements in status estimates, or work to expand access to long-running existing surveys that have yet to be fully utilized in fisheries assessment [@maureaud2020]. -->

```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error (RMSE) of *B/B~MSY~* in the most recent 5 years of data available for each model fit. Distribution is full posterior probability distribution. Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th. Change is relative to the mean performance of a catch-only heuristic model."}

voi_plot <- b_voi_plot + 
  geom_vline(aes(xintercept = 0), alpha = 0.75)

voi_plot

ggsave(here("documents","figs","voi-plot.png"),voi_plot, width = 8, height = 6)

```

#  Discussion {.unnumbered}

<!-- XX somewhere in here make clear that you're not learning anything from the catch data XX -->

<!-- XX Acknolwedge the need for action the the frustration of being told "get more data" XX -->

<!-- Global-level assessment are critical for guiding management agendas for the world's oceans. Despite this need, and despite advances in stock assessment methods, our understanding of the world's fisheries remains murky in many parts of the world. Numerous efforts have sought to resolve this problem through development of new modeling methods. These studies have ranged from simple heuristic algorithms to superensemble modeling, but while these studies have advanced our knowledge and motivated important conversations about the state of unassessed fisheries, each has demonstrable shortcomings as an answer for global fisheries assessment. However, a growing number of broadly (if not globally) distributed datasets have become available recently, leading us to ask whether using these data to move beyond "catch-only" models for global assessment would produce more reliable outcomes. While we demonstrate that these data can provide value above and beyond the catch histories alone, addition of these data alone does not substantially clarify the picture of global fisheries status.   -->

<!-- Our results show that estimates of stock status based on catch histories alone stem almost entirely from underlying assumptions, not the information content of the catches, producing results that are inconsistent with best-available knowledge of the state of assessed fish populations in RLSADB -->

Global-level assessments of fish populations are critical for guiding management agendas for the world's oceans, and tracking indicators such as the United Nations Sustainable Development Goals. The hope of efforts using catch-only models to estimate the status of global unassessed fisheries is fundamentally that we can learn something meaningful about the state of a fished population simply by knowing something about its catch history and life history. While in some cases the addition of globally available data such a FMI scores, SAR values, or effort reconstructions, provided value above and beyond catch histories alone (Fig.\@ref(fig:cs-plot)), at the global level models fit using each of the available datasets, besides the RLSADB-derived indices, generally produced biased and imprecise estimates of fish stock status (Table.\@ref(tab:perf-tab)). The simplicity and low data requirements of catch-only models are understandably appealing to many users, but our results highlight that without high-quality local data these methods can provide highly imprecise and biased results. Broad and uncritical application of these methods that frequently result in incorrect classification of stock status can be detrimental to both ecosystems and livelihoods.

<!-- That use of none of the datasets collected here can achieve that, and that our tests fitting directly to abundance indices from RLSADB itself suggests that model misspecification is not the primary cause of poor performance, tells us that improvements in estimates of global stock status will likely depend on improvements in the quality and use of data themselves.  -->

Our claim is that achieving substantial progress in assessing the state of global fisheries will require improved data and capacity building. This claim cannot be definitively proven; it is possible that some yet-to-be discovered model will provide a reliable means of dramatically improving the accuracy of estimates produced by catch-based models using current broadly-available datasets. Basic logic tells us though that such a model would have to depend on strong assumptions, new data, or empirical estimation. Catch is a function of catchability, effort, and biomass, so given just catch we have one equation and three unknowns. Structurally separating out biomass (and subsequently estimating reference points etc.) requires then either data or strong assumptions on these other variables across a sufficiently representative timespan. Alternatively, empirical relationships could be developed linking attributes of catch histories to observed estimates of stock status [@costello2012b;@thorson2012a], or through identifying a structural model for the evolution of fishing mortality in managed and unmanaged fisheries [@thorson2013]. However, these methods require that these empirical relationships both exist and are reliably preserved in fisheries that lack formal assessments. Numerous groups made up of highly skilled scientists have attempted to resolve these fundamental challenges in various ways, but external efforts to validate the performance of these methods have consistently found them to perform generally poorly when applied in bulk [@free2020;@bouch2021;@pons2020;@sharma2021]. This summary is not intended as a critique of those past efforts, but a reflection on the limited amount of broadly applicable predictive power that exists within catch histories alone. Methods tested in those papers have been demonstrated to perform relatively well when given specific data or priors informative to specific fisheries. We revisited this task in this paper, and found that even when augmented with other broadly available datasets, we were unable to produce consistently accurate estimates of stock status around the globe. For these reasons we believe that improvements to our understanding of the state of global unassessed fisheries will come from targeted use of existing but underutilized data, and expanding the collection of high-priority data around the globe, rather than the development of new modeling methods alone.

What quality of assessment is needed and what constitutes a meaningful improvement in assessment quality depends on the needs of those using the assessment outputs. It may be that for particular regions, species, or uses the results presented here or in other past global analyses are sufficiently accurate. Catch-only based estimates of measures such as MSY are likely to be much more robust than reference points such as *B/B~MSY~* [@martell2013]. Where catch-only methods are used for estimates of *B/B~MSY~* or *F/F~MSY~*, it is critical that users are aware of the broadly demonstrated poor performance of these methods in the absence of highly reliable fishery-specific information. In some instances using the data presented here did provide some improvement over use of catch-only style assessment methods; the difficulty comes in attempting to apply data types uniformly across the globe. While it is unreasonable to expect models based solely on global-scale data to be able to perform as well as detailed stock assessments reported in RLSADB, or that data-limited methods would perform well for every individual stock, our hope would be that a data-limited approach based on globally available data sources would be able to correctly capture general patterns in stock status in time and space. The overall poor performance of the models tested here, in terms of estimating both *B/B~MSY~* and *F/F~MSY~*, shows that improvements in estimates of global stock status depend on improvements in the quality and use of data themselves.

Our fits to the RLSADB data provide a useful diagnostic of the degree to which model mispecification might explain the poor performance of catch-only models. Catch-only models generally employ some form of biomass dynamics model, such as the Schaefer surplus production model [@schaefer1954]. These models are generally much simpler than the typically age-structured population models underpinning the assessments in the RLSADB. While we found that fitting the Pella-Tomlinson model employed in sraplus to an index of abundance drawn directly from the RLSADB provided much improved and reasonably accurate estimates of stock status relative to catch-only approaches (Table.\@ref(tab:perf-tab)), even with the index of abundance sraplus still had an average MAPE of `r scales::percent(perf$mape[perf$'Data Used' == "RLSADB Index"])`. This discrepancy may provide a measure of relatively insurmountable error resulting from model misspecification that inclusion of additional data may not be able to overcome.  For example, many assessments in the RLSADB calculate *B/B~MSY~* based on spawning stock biomass. However, surplus production models condition the size and state of the population on the catch, which is by definition a measure of the vulnerable biomass. Vulnerable biomass may not reliably track spawning biomass in highly selective fisheries, potentially resulting in model biases as the surplus production model is not capable of separately tracking spawning and vulnerable biomass.

We chose to test the performance of methods against estimated values reported in RLSADB. A reasonable critique of this choice is that unassessed stocks, on which these methods would actually be used, are likely to have vastly different dynamics than the heavily managed fish populations represented in RLSADB, in addition to the errors in the RLSADB estimates themselves. For this reality to change our results though would require that the methods tested here have massively lower bias and higher accuracy for unassessed fisheries than RLSADB stocks. @free2020's simulation based results of these same types of models suggests that this is unlikely to be true. However, while many assessed fisheries have mechanisms to actively manage catches, catch histories in unassessed and unmanaged fisheries may be more likely to reflect underlying stock status than those in relatively well managed fisheries, though any predictive relationship will still depend on a host of factors such as changes in effort. This could be empirically tested by considering the performance of catch-only models on data from earlier in the history of fisheries in the RLSADB, though we did not do so here as we do not have historic estimates of FMI or SAR values.

<!-- Our estimates of global stock status vary wildly depending on which source of data we use. Why not simply pick the best source of data, or use all the available data, and use that to update our estimate of global stock status, under the grounds that that is the best available information? Unfortunately, it is not clear that we can evaluate which source of data provides the best fit; examination of each shows strengths and weakness that are not easily comparable (Fig.\@ref(fig:mape-map)). Globally available effort data from @rousseau2019 allows us to create local indices of abundance based on catch histories but carries a host of critical assumptions, namely that the same effort history applies to all fisheries in the region, or country if country level effort data are used. These CPUE indices are also extremely sensitive to assumptions around rates of change in catchability. The fisheries management index has some ability in predicting stock status (Fig.XX), but out-of-sample testing reveals limitations. Swept-area ratios have a strong relationship with fishing mortality rates, but only for comparable fisheries; it is unclear how good a proxy this measure is for non-demersal stocks. None of these data, or combinations thereof, was able to satisfactorily match the FAO's SOFIA assessments, or the assessments presented in RLSADB (XX RLSADB matches? need to add an analysis: match stocks from SOFIA to RLSADB stocks where possible, plot comparison).  -->

Our results do not imply that the kinds of broadly available data presented here are not valuable under the right conditions. The FMI and SAR based priors are an improvement over catch-only models in applicable situations (i.e. those that sufficiently resemble the data on which the regressions were trained, Fig.\@ref(fig:cs-plot)). Effort data such as those reconstructed by @rousseau2019 can help distinguish between regions with similar catch histories but different large-scale effort trajectories, and may be quite useful as indices of abundance for areas with good knowledge of rates of evolution of fishing technology and a broadly selective fishing fleet. Despite not adding a great deal in terms of performance at the global scale, swept-area-ratio was the strongest predictor of *F/F~MSY~* of any of the datasets we explore on an individual stock basis, with a Bayesian R^2^ value of `r round(mean(rstanarm::bayes_R2(sraplus::sar_models$fit[[3]])),2)` (see SI). But, we must simultaneously consider data quality and resolution: applying one SAR value to all stocks in a region, even if that SAR value can provide valuable information for a subset of fisheries, causes inaccurate estimates of stock status when applied too broadly. Our analysis does not show that the data considered here are without value, but that attempting to indiscriminately apply these data to all areas of the globe results in meaningfully incorrect estimates of stock status for regions whose nature does not match the assumptions needed to apply these data sources. We found that performance of different data sources varied widely both within and among regions (Fig.\@ref(fig:cs-plot), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). Some of this variation is likely simply due to low sample sizes of assessed stocks reported in the RLSADB in some regions. But, other explanations for differences in performance among stocks and regions may be an interesting area for future research. For example, it may be that stocks in some regions are more suitable to being represented by surplus production models than others.

Our value of information analysis also shows the high utility of having access to even a recent snapshot of *F/F~MSY~* (Fig.\@ref(fig:voi-plot)). Swept area ratios, Fisheries Management Index scores, or other similar metrics can be used to construct fishery-specific priors on fishing mortality rates, though care must be taken in applying them at the appropriate spatial resolution. Another avenue would be to prioritize the development of a global repository for length and age composition data. Given appropriate conditions, these length measurements can be used to estimate local fishing mortality rates [@hordyk2016; @prince2019; @rudd2017]. While length-based assessments come with a host of assumptions and potential pitfalls, properly implemented in some fisheries with appropriate life histories these methods may provide an overlooked source of information on fisheries at a global scale, at least as an improvement over relying on catch-only or regional proxies alone. Such a database could be used to construct stock or stock complex specific priors on fishing mortality for particular regions around the globe, which could meaningfully improve our understanding of global fisheries, particularly when paired with catch data and where possible indices of abundance [@thorson2015f; @rudd2017].

<!-- Using catch data alone does not reveal about recent stock status beyond what is encoded in our priors and functional forms. So if we had good estimates of B/B~MSY~ our job would already be done without any need to run something like `sraplus` (except to derive other quantities of interest). Given the demonstrated value of F/F~MSY~ values in improving estimates of stock status, and the increased feasibility in estimating F/F~MSY~ at a local scale relative to B/B~MSY~, how might we go about obtaining estimates of fishing mortality rates at a global scale? We have demonstrated one potential approach here; fisheries management index and swept area ratio values both provide reasonable predictions of F/F~MSY~ for individual fisheries (see Supplemental Information). While assigning mean FMI and SAR values at the taxonomic and regional level as we did in Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map) did not perform particularly well, efforts to assign FMI and SAR values to individual stocks or appropriate stock complexes might be more fruitful. -->

We must also prioritize collection and curation of fish population survey data worldwide. Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://www.FishStats.org) ( www.FishStats.org). Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have stock assessments. Effort reconstructions such as those utilized here may help create fishery-dependent abundance indices in some instances, and going forward datasets such as those compiled by [Global Fishing Watch](https://globalfishingwatch.org/) (https://globalfishingwatch.org/) in combination with the reconstruction approaches of @rousseau2019 might allow us to construct and use timeseries of fishing effort specific to particular areas, fleets, and species complexes. However, our value of information exercise indicates that we may have to wait many years for new surveys to provide substantial improvements in status estimates (Fig.\@ref(fig:voi-plot)).

Expanded training of fisheries scientists around the globe is another critical need. Even were we to dramatically expand the amount and types of data available for global assessment, individual fisheries and regions will need to make informed decisions about which sources of data may be applicable and which not, and to critically evaluate the results of any model based on local expertise. This is why stock assessments even in data-rich fisheries are not an automated process; the real challenge is often not in fitting a model to data but in understanding how best to use the data and the quality and limitations of the model used. Empowering a global network of fisheries scientists through training and peer-support would help local experts make the most of available data, ensure the reliability of newly collected data, and improve the interpretation of assessment results. We may also need to acknowledge that in some fisheries reliable estimates of stock status relative to MSY based reference points are simply not possible (or perhaps desirable), and instead rely on more precautionary or empirical management management measures such as spatial closures, size restrictions, and indicator based harvest strategies (ideally tested through management strategy evaluation) [@prince2019;@dowling2015;@fulton2016].

The coming decades are a critical time for the future of fisheries and ocean health. Achieving the United Nations Sustainable Development Goal 14 for the conservation and sustainable use of the world's oceans depends on our ability to effectively assess the status of fish stocks around the world. The RAM Legacy Stock Assessment Database combined with the FAO's expert elicitation of status for select stocks have dramatically improved our understanding of global fisheries in recent years. However, this process still leaves a substantial number of fisheries and proportion of global catch unassessed. Numerous catch-based data-limited approaches have attempted to fill that gap, and while these efforts have advanced our knowledge and interest in unassessed fisheries, none have yet been able to provide a solution to this problem which has proven to be unbiased and sufficiently precise at a global or regional level.

<!-- Our results show that improving estimates of global stock status depends on investing in an improved and expanded global network of fishery data and fisheries scientists. -->

The lack of strong information on stock status within catch histories alone means that differences in models and assumptions between catch-based assessment efforts can produce starkly contrasting conclusions on global stock status, leading to debates that are inconclusive as they are inherently driven by assumptions. The FAO is leading efforts to increase technical capacity and monitoring and evaluation infrastructure to improve fisheries management in places with limited data. Such projects stand to provide a better picture of fishery status at global and local scales, furthering our ability to meet the UN SDG targets. Our results emphasize the urgency and rationale for building the infrastructure and capacity that can lead to better marine resource management globally [@costello2020] Achieving meaningful improvements in the assessment and management of global unassessed fisheries will depend on expanded collection of targeted data types, active management, and local capacity building.

<!-- The need is to develop better datasets and approaches on the ground that can lead to more conclusive inferences about the state of global unassessed fisheries. -->

# Data Availability Statement {.unnumbered}

All data and materials needed to reproduce our results are publicly available or queried by code available at https://github.com/DanOvando/assessing-global-fisheries.

# Acknowledgements {.unnumbered}

Funding for this work was provided by the Food and Agriculture Organization of the United Nations (FAO). We thank participants of the FAO "Methods for Global Assessment" workshop held in Rome, Italy February 2019 for helpful feedback on this project. We also thank A. Hordyk and one anonymous reviewer for their helpful comments. The scientific results and conclusions, as well as any views or opinions expressed herein, are those of the author(s) and do not necessarily reflect those of the FAO, NOAA or the Department of Commerce.

# References {.unnumbered}

::: {#refs}
:::

\newpage

# Tables {.unnumbered}

+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Data Source                                 | Short Name       | Data Use                                                                 | Caveats                                                                                                         |
+=============================================+==================+==========================================================================+=================================================================================================================+
| Catch data [@fao2020]                       | catches          | Priors on stock status, scaling of population size, exploitation history | Heuristics or regressions used to translate shape of catch history into priors on stock status                  |
+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Fisheries Management Index [@melnychuk2017] | FMI              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Swept Area Ratio [@amoroso2018]             | SAR              | Priors on most recent *F/F~MSY~* values                                    | Priors produced by regression trained on data from RAM Legacy Stock Assessment Database                         |
+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
| Reconstructed effort data [@rousseau2019]  | effort           | Combined with catch data to create an index of abundance                 | Total reconstructed effort across all sectors. Assumed rate of technology creep reported in individual sections |
+---------------------------------------------+------------------+--------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------+
:(\#tab:data) Data sources included across model fits.

```{r dat-desc}
dat_description <- tribble(~`Data Name`, ~Description,
        "RLSADB Index", "Fit to abundance index from RLSADB",
        "SAR", "Prior on terminal F/Fmsy set by regional swept area ratio",
        "FMI", "Prior on terminal F/Fmsy set by regional fisheries management index scores",
        "Efective CPUE", "Fit to CPUE index created from RLSADB catch and regional effort index. 2.6% technology creep",
        "Effective CPUE+", "Fit to CPUE index created from RLSADB catch and regional effort index with priors informed by SAR and FMI. 2.6% tech. creep",
        "Nominal CPUE", "Fit to CPUE index created from RLSADB catch and regional effort index. 0% tech. creep",
         "Nominal CPUE+", "Fit to CPUE index created from RLSADB catch and regional effort index with priors informed by SAR and FMI. 0% tech. creep",
        "Guess", "Priors on terminal B/Bmsy randomly sampled from 0.4,1.0,1.6")

knitr::kable(dat_description,
             caption = "Data sources used for terminal stock status estimate",
             format = "pipe") %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

```{r perf-tab}

knitr::kable(perf %>% rename(MPE = mpe, MAPE = mape, Accuracy = accuracy), digits = 2,
             caption = "Global performance statistics in the most recent year available of models using different sources of data. MPE = median percent error (bias), MAPE = median absolute percent error (error), Accuracy = percent of times that stocks were classified to the correct FAO status bin (underfished, maximally sustainably fished, overfished). Performance is judged relative to *B/B~MSY~* reported values in RAM Legacy Stock Assessment Database.",
              format = "pipe")

```

\newpage
# Figure Legends  {.unnumbered}

Figure 1: RLSADB values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], catch heuristics, priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort (Effort) trends assuming a rate of technological increase of 2.6%. Each point is a stock in the RLSADB. Black dashed line shows the 1:1 relationship. Text displys root mean squared error (RMSE) of each panel.

Figure 2: Median percent error (MPE, predicted relative to observed) in most recent *B/B~MSY~* by FAO statistical area from different data sources. Data source panels are ordered in ascending (starting from top left) total bias at the FAO region level. RLSADB Index refers to catch and abundance index drawn from RLSADB. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.

Figure 3: Median absolute percent error (MAPE) in most recent *B/B~MSY~* by FAO statistical area from different data sources. Data source panels are ordered in ascending (starting from top left) mean MAPE at the FAO region level. RLSADB Index refers to catch and abundance index drawn from RLSADB. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.

Figure 4: Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources. Data source panels are ordered in descending (starting from top left) mean accuracy at the FAO region level. RLSADB Index refers to catch and abundance index drawn from RLSADB. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for 'effective' a 2.6% technology creep is assumed. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assigns a random recent *B/B~MSY~* of 0.4,1, or 1.6.

Figure 5: Posterior probability distributions of estimated effect of different data types on root mean squared error (RMSE) of *B/B~MSY~* in the most recent 5 years of data available for each model fit. Distribution is full posterior probability distribution. Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th. Change is relative to the mean performance of a catch-only heuristic model.

<!-- \newpage -->

<!-- # Supporting Information {.unnumbered} -->

<!-- ```{r si, child = 'ovando-etal-assessing-global-fisheries-si.Rmd'} -->
<!-- ``` -->